---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(cluster)
library(factoextra)
library(dendextend)
library(GGally)
```

```{r}
mall <- read_csv("data/mall_customers.csv") %>% 
  janitor::clean_names()

glimpse(mall)
```
*Some visualisations*

```{r}
mall %>% 
  ggplot() +
  aes(x = gender, fill = gender) + 
  geom_bar()
```
```{r}
mall %>% 
  ggplot() + 
  aes(x = gender, y = annual_income_k, fill = gender) + 
  geom_col() 
```
```{r}
mall %>% 
  ggplot() + 
  aes(x = gender, y = spending_score_1_100, fill = gender) + 
  geom_col() 
```

```{r}
mall %>% 
  ggplot() + 
  aes(x = age, y = spending_score_1_100, fill = gender) + 
  geom_col() 
```

There are more women than men in this sample.
The women collectively have a higher income than the men.
Collectively, the women have a higher spending_score than the men also. 


*K means clustering*

Plotting the data - maybe there are 4 or 5 clusters.

```{r}
mall %>% 
  ggplot() +
  aes(x = spending_score_1_100, y = annual_income_k) +
  geom_point()
```

Use numeric methods to determine potential number of clusters.

```{r}
# scale the numeric mall data 
mall_scale <- mall %>%
  mutate_if(is.numeric, scale)
```

1. Elbow

```{r}
fviz_nbclust(mall_scale %>% 
            select(spending_score_1_100, annual_income_k), 
            FUNcluster = kmeans, 
            method = "wss", 
            nstart = 20)
```
Maybe 5 clusters?


2. Silhouette

```{r}
fviz_nbclust(mall_scale %>% 
            select(spending_score_1_100, annual_income_k), 
            FUNcluster = kmeans, 
            method = "silhouette", 
            nstart = 20)
```

Again, looks like 5 clusters.


3. Gap Statistic

```{r}
fviz_nbclust(mall_scale %>% 
            select(spending_score_1_100, annual_income_k), 
            FUNcluster = kmeans, 
            method = "gap_stat", 
            nstart = 20)
```

This recommends 1 cluster which isn't correct.

Let's go with 5 clusters. 

```{r}
clusterings <- kmeans(
  mall_scale %>% 
    select(spending_score_1_100, annual_income_k), 
  centers = 5, 
  nstart = 20
)
```

```{r}
augment(clusterings, mall) %>% 
  ggplot() +
  aes(x = spending_score_1_100, 
      y = annual_income_k, 
      colour = .cluster) +
  geom_point() + 
  # theme(legend.position = "") +
  labs(
    x = "spending score", 
    y = "annual income", 
    title = "Shopper clusters", 
    
  )
```

The clustering seems a very good fit.

Here are the mean scores for each of the numerical variables in the `mall` data, grouped by the cluster number.

```{r}
aggregate(mall, by=list(cluster=clusterings$cluster), mean) %>% 
  select(-customer_id, -gender)
```

1 = older, high income, high spending
2 = older, low income, low spending
3 = younger, low income, high spending score
4 = older, high inome, low spending score
5 = middle age, high income, high spending score
