---
title: "wk2_day_01_homework"
output: html_notebook
---

# Q1 - Read in the data

Read in the original data
```{r}
card_base <- read_csv('data/CardBase.csv')
customer_base <- read_csv('data/CustomerBase.csv')
fraud_base <- read_csv('data/FraudBase.csv')
transaction_base <- read_csv('data/TransactionBase.csv')
```

load janitor
```{r}
library(janitor)
```

Clean the column names. 
```{r}
card_base <- clean_names(card_base)
customer_base <- clean_names(customer_base)
fraud_base <- clean_names(fraud_base)
transaction_base <- clean_names(transaction_base)
```


# Q2

```{r}
nrow(card_base)
```

Given that there are 500 entries in the card_base table, there will be 500 rows when left_joined to the customer_base table. 

```{r}
left_join(card_base, customer_base, by = "cust_id")
```

# Q3

Entries in fraud and transaction tables
```{r}
nrow(fraud_base)
nrow(transaction_base)
```

Full join between fraud_base and transaction_base tables
```{r}
full_join(fraud_base, transaction_base, by = "transaction_id")
```

The resulting table from the full join has the same number of entries as the transaction table, which means that the fraudulent transactions were already recorded within the transaction table. 


# Q4

Semi-join is required here. 

```{r}
semi_join(card_base, transaction_base, by = c('card_number' = 'credit_card_id'))
```

500 rows were returned. 

Number of rows in card_base table is also 500. 

```{r}
nrow(card_base)
```


# Q5

Read in hat_observations
```{r}
hat_obs <- read_csv("data/hat_observations.csv")
```

Separate observation column. 
```{r}
hat_obs <- hat_obs %>% 
  separate(observation, c("hat_color", "hat_type"), sep = "\\,")

hat_obs
```

# Q6

Join day, month and year in to one column. 
```{r}
hat_obs <- hat_obs %>% 
  unite(date, c("day", "month", "year"), sep = "-")

hat_obs
```


The date when the most berets were observed. 

```{r}
hat_obs %>%
  group_by(date, hat_type) %>% 
  summarise(hat_count = sum(observation_count)) %>% 
  filter(hat_type == "beret") %>% 
  arrange(desc(hat_count)) %>% 
  head(1)
```


