---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(skimr)
```

```{r}
# load the data
red_wine <- read_csv("data/wine_quality_red.csv")
white_wine <- read_csv("data/wine_quality_white.csv")
```


*Inspect the data*

```{r}
names(red_wine)
```
```{r}
names(white_wine)
```
Names are the same in both data sets. 

```{r}
glimpse(red_wine)
```
```{r}
glimpse(white_wine)
```
Both data sets have the same variable names and data types. 

We should join the two tables with `bind_rows()`. 

```{r}
wine <- bind_rows(red_wine, white_wine)
```


*Alias*

Are there any aliases in the dataset?

```{r}
alias(quality ~., data = wine)
```
No aliases. 


*Categorical variables & tidy-up*

```{r}
# NAs in the dataset
colSums(is.na(wine))
```
Only 2 `NA` observations from ~6,500. 

Remove `region` as a variable. It doesn't characterise each of the wines and is too generic to be of value. 

```{r}
wine_trim <- wine %>% 
  select(-region, -wine_id) %>% 
  drop_na() # remove NAs
```

Use `skim()` function to examine dataset. 

```{r}
wine_trim %>% 
  skim() %>% 
  view()
```

- `p_h` has values in excess of 14 and heavily left-skewed
- `quality` values are within the expected range
- `alcohol` has values in excess of 100abv (which cannot be correct) and is heavily right-skewed.
- all other variables are right-skewed

Plot `p_h`:

```{r}
wine_trim %>% 
  ggplot() + 
  aes(x = p_h) +
  geom_bar()
```
Something is amiss with the `p_h` data. It should be in the range 0 - 14 but the vast majority of the values are between 300 to 400. Consider log transformation of existing variable.

For `quality`, cannot convert to data type `factor` as the dependent variable cannot be categorical.

All the numerical variables are skewed (most to the right) and should be log transformed for linear regression. 

Variable `citric acid` has zero values which produce `inf` results when log transformed so have added small positive value (1) to each value prior to log transformation. 

```{r}
wine_trim <- wine_trim %>% 
  mutate(log_fixed_acidity = log(fixed_acidity)) %>% 
  mutate(log_volatile_acidity = log(volatile_acidity)) %>% 
  mutate(log_citric_acid = log(citric_acid + 1)) %>% 
  mutate(log_residual_sugar = log(residual_sugar)) %>% 
  mutate(log_chlorides = log(chlorides)) %>% 
  mutate(log_free_sulfur_dioxide = log(free_sulfur_dioxide)) %>% 
  mutate(log_total_sulfur_dioxide = log(total_sulfur_dioxide)) %>% 
  mutate(log_density = log(density)) %>% 
  mutate(log_p_h = log(p_h)) %>% 
  mutate(log_sulphates = log(sulphates)) %>% 
  mutate(log_alcohol = log(alcohol)) %>% 
  select(-c(fixed_acidity, volatile_acidity, citric_acid, residual_sugar, chlorides,
            free_sulfur_dioxide, total_sulfur_dioxide, sulphates, density, p_h, alcohol))
```

View histogram of variables with skimr:

```{r}
wine_trim %>% 
  skim() %>% 
  view()
```

*Issues:*

Distribution of variables after transformation is much improved although `log_density`, `log_ph` and `log_alcohol` all remain badly skewed.  Have tried other transformations of these variables (reciprocal and square root) but to no avail. Have decided to remove them before building the model. 


```{r}
# remove skewed variables
wine_trim <- wine_trim %>% 
  select(-c(log_density, log_p_h, log_alcohol))
```


*Model build*

```{r, message=FALSE, warning=FALSE}
ggpairs(wine_trim)
```

*First predictor*

There appears to be (an albeit weak) correlation between `quality` and `log_volatile_acidity` and `log_chlorides`:

mod1a - `log_volatile_acidity`
mod1b - `log_chlorides`

```{r}
mod1a <- lm(quality ~ log_volatile_acidity, data = wine_trim)
mod1b <- lm(quality ~ log_chlorides, data = wine_trim)
```

_diagnostic plots_

```{r}
autoplot(mod1a)
```
residual vs fitted - looks fine as the residuals appear randomly scattered around zero
normal Q-Q - again looks fine as residuals stick close to line and so appear normally distributed 
scale/location - fixed width band plot of residuals which is fine

```{r}
autoplot(mod1b)
```
again, these plots are similar to `mod1a` plots and appear fine. 

_summary statistics_

```{r}
summary(mod1a)
```
intercept = 5.242
r^2 = 0.05249
rse = 0.9089
although statistically significant, `log_volatile_acidity` explains very little of the variation in quality

```{r}
summary(mod1b)
```
intercept = 4.292
r^2 = 0.05541
rse = 0.9075
like mod1a, `log_chlorides` is statistically significant but explains very little of the variation in `quality`.

Both predictors have poor r^2 values but the better of the two is `mod1b` so we will choose `log_chlorides` as the first predictor for the model. 

Model: quality = 4.29 + (-0.51 * `log_chlorides`)


*Second predictor*


```{r}
# remove quality and log_chlorides from model
wine_remaining_resid <- wine_trim %>% 
  add_residuals(mod1b) %>% 
  select(-c(quality, log_chlorides))
```

```{r, warning=FALSE, message=FALSE}
# examine relationship between remaining predictors
ggpairs(wine_remaining_resid)
```

Again, the correlation between the residuals and the remaining predictors is weak. The best of the bunch are:

mod2a - `log_chlorides` + `log_volatile_acidity`
mod2b - `log_chlorides` + `log_sulphates`

```{r}
mod2a <- lm(quality ~ log_chlorides + log_volatile_acidity, data = wine_trim)
mod2b <- lm(quality ~ log_chlorides + log_sulphates, data = wine_trim)
```

_diagnostic plots_

```{r}
autoplot(mod2a)
```
Very similar to the original plots for mod1a and mod1b, the diagnostic plots look fine


```{r}
autoplot(mod2b)
```
Again, the plots look fine. 

_summary statistics_

```{r}
summary(mod2a)
```
intercept = 4.34
r^2 = 0.0751
rse = 0.8981
like the previous predictors, whilst `log_volatile_acidity` is statistically significant, the r^2 value is very small and is a poor predictor. 

```{r}
summary(mod2b)
```
intercept = 4.28
r^2 = 0.0764
rse = 0.8975
as above, the `log_sulphates` predictor is statistically significant but with such a low r^2 number offers little to the model. That said, it provides more predictive power than `mod2a` so we choose `mod2b` to be the new model. 

Model: quality = 4.29 + (-0.51 * `log_chlorides`)

Model: quality = 4.28 + (-0.64 * `log_chlorides`) + (0.571 * `log_sulphates`)


*Third predictor*

```{r}
# remove log_sulphates from model
wine_remaining_resid <- wine_trim %>% 
  add_residuals(mod2b) %>% 
  select(-c(quality, log_chlorides, log_sulphates))
```

```{r, warning=FALSE, message=FALSE}
# examine relationship between remaining predictors
ggpairs(wine_remaining_resid)
```

Yet again, the correlation between the residuals and the remaining predictors is weak. The best of the bunch are:

mod3a - `log_chlorides` + `log_sulphates` + `log_volatile_acidity`
mod3b - `log_chlorides` + `log_sulphates` + `log_total_sulfur_dioxide`

```{r}
mod3a <- lm(quality ~ log_chlorides + log_sulphates + log_volatile_acidity, data = wine_trim)
mod3b <- lm(quality ~ log_chlorides + log_sulphates + log_total_sulfur_dioxide, data = wine_trim)
```

_diagnostic plots_

```{r}
autoplot(mod3a)
```
As with the previous plots, these look fine. 


```{r}
autoplot(mod3b)
```
residual vs fitted: slight smile on the line but nothing major. 
All other plots look fine. 

_summary statistics_

```{r}
summary(mod3a)
```
intercept = 4.328
r^2 = 0.099
rse = 0.886
like all the others before, this predictor is statistically significant but offer little predictive power

```{r}
summary(mod3b)
```
intercept = 4.565
r^2 = 0.083
rse = 0.8945
yet again, whilst statistically significant, the predictor adds little to the model. 


Although dire, `mod3a` offers the best results so we adopt `log_volatile_acidity` in to the model. 

Model: quality = 4.328 + (-0.493 * `log_chlorides`) + (0.621 * `log_sulphates`) 
                       + (-0.363 * `log_volatile_acidity`)

