---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
```

```{r}
houses <- read_csv("data/kc_house_data.csv")
```


# Question 1

```{r}
head(houses)
```

```{r}
# create logical variables for waterfront and renovated
houses_tidy <- houses %>%   
mutate(waterfront = as.logical(waterfront)) %>%  
mutate(renovated = as.logical(if_else(yr_renovated > 0, 1, 0)))
```

```{r}
# remove variables
houses_tidy <- houses_tidy %>% 
  select(-c(id, date, sqft_living15, sqft_lot15, zipcode, yr_renovated)) 
```

Leave `view`, `condition` and `grade` as they are. All are categorical ordinal variables with an implicit order to them - the higher the number the better the category. 


# Question 2

```{r}
# test for alias
alias(lm(price ~ ., data = houses_tidy))
```
```{r}
# remove alias
houses_tidy <- houses_tidy %>% 
  select(-sqft_basement)
```

```{r}
# test for alias again
alias(lm(price ~ ., data = houses_tidy))
```

# Question 3

```{r}
houses_tidy_numeric <- houses_tidy %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses_tidy %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- houses_tidy$price
```

```{r, message=FALSE, warning=FALSE}
ggpairs(houses_tidy_numeric)
```
*Numeric variables:*

High correlation pairs:
- price & sqft_living (0.702)
- price & grade (0.667)
- price & sqft_above (0.606)
- price & bathrooms (0.525)
- price & bedrooms (0.308)



```{r, message=FALSE, warning=FALSE}
ggpairs(houses_tidy_nonnumeric)
```
Variables of interest:
- `waterfront` looks to have variation in mean between the 2 categories


*Predictor 1*

```{r}
mod1a <- lm(formula = price ~ sqft_living, data = houses_tidy)
mod1b <- lm(formula = price ~ waterfront, data = houses_tidy)
```

_mod1a_

```{r}
autoplot(mod1a)
```
```{r}
summary(mod1a)
```
r^2 = 0.4929
res = 261500
diagnostic plots look ok altho skewness at high end on Normal Q-Q
`sqft_living` is a statistically significant predictor

_mod1b_

```{r}
autoplot(mod1b)
```
```{r}
summary(mod1b)
```
r^2 = 0.07095
rse = 353900
diagnostic plots look ok but there is skewness at the high end on Normal Q-Q
`waterfront` is a statistically significant predictor

-------------------------------------------------------------------------------

Given the r^2 number, first predictor is `sqft_living`. 

model: price =  -43580.743 + (280.624 * `sqft_living`)


*Predictor 2*

```{r}
# remove variables included in the model
houses_remaining_resid_numeric <- houses_tidy_numeric %>% 
  add_residuals(model = mod1a) %>% 
  select(-c(price, sqft_living))
```

```{r, message=FALSE, warning=FALSE}
ggpairs(houses_remaining_resid_numeric)
```
*Numeric variables:*

High correlation pairs:
- residual & lat (0.379)
- residual & view (0.277)


Hmmm, let's add `lat` from the numeric predictors.

```{r}
mod2a <- lm(formula = price ~ sqft_living + lat, data = houses_tidy)
mod2b <- lm(formula = price ~ sqft_living + view, data = houses_tidy)
```


_mod2a_

```{r}
autoplot(mod2a)
```
```{r}
summary(mod2a)
```
r^2 = 0.566
rse = 241900
diagnostic plots look ok but there is skewness at the high end on Normal Q-Q
`lat` is statistically significant as a predictor

_mod2b_

```{r}
autoplot(mod2b)
```
```{r}
summary(mod2b)
```
r^2 = 0.5353
rse = 250300
diagnostic plots look ok but there is skewness at the high end on Normal Q-Q
`view` is statistically significant as a predictor

-------------------------------------------------------------------------------

The r^2 number for 'lat' is marginally higher than for `view`, so `lat` becomes our second predictor. 

The r^2 number has risen from 0.4929 (`sqft_living`) to 0.566 with the addition of `lat`. 

The rse has fallen from 261500 to 241900. 

The regression equation is now:

price =  -34160000 + (274.9 * `sqft_living`) + (717700 * `lat`)



