---
title: "Wk1 Weekend homework"
output: html_notebook
---

```{r}
# install packages  
library(tidyverse)
library(janitor)
```

```{r}
# load books dataset 
books_data <- read_csv("data/books.csv")
```

```{r}
# examine the data
glimpse(books_data)
view(books_data)
```

Q1: Which publisher has published the most books?

```{r}
books_data %>% 
  group_by(publisher) %>% 
  summarise(books_published = n()) %>% 
  slice_max(books_published, n = 5) # top 5 shown for context
```


Q2: The top 5 rated books?

```{r}
books_data %>% 
  select(title, average_rating) %>% 
  slice_max(average_rating, n = 5, with_ties = FALSE)
```


Q3: Which publisher has the best rating?

Question about rounding - the avg rating for Academia press should be 4.5 not 5. Why is it not displayed as such. 

```{r}
books_data %>% 
  group_by(publisher) %>% 
  summarise(avg_rating = mean(average_rating)) %>% 
  arrange(desc(avg_rating)) %>% 
  head(5) # show top 5
```

Compare result to using slice_max instead of `arrange()` and the `head()`:

```{r}
books_data %>% 
  group_by(publisher) %>% 
  summarise(avg_rating = mean(average_rating)) %>% 
  slice_max(avg_rating, n = 5, with_ties = FALSE)
```



Q4:  Which author has the highest average page count?

```{r}
books_data %>% 
  group_by(authors) %>% 
  summarise(avg_page_count = mean(num_pages)) %>% 
  slice_max(avg_page_count, n = 5, with_ties = FALSE)
```

Q5: How many books are anonymous?

```{r}
books_data %>% 
  filter(authors == "Anonymous") %>% 
  summarise(authors, book_count = n())
```

Q6: The top 5 languages used?

```{r}
books_data %>% 
  group_by(language_code) %>% 
  summarise(count = n()) %>% 
  slice_max(count, n = 5, with_ties = FALSE)
```

Q7: The most reviewed books?

```{r}
books_data %>% 
  group_by(title) %>% 
  summarise(reviews = sum(text_reviews_count)) %>% 
  slice_max(reviews, n = 5, with_ties = FALSE)
```


Q8: The most published authors?

```{r}
books_data %>% 
  group_by(authors) %>% 
  summarise(published_books = n()) %>% 
  slice_max(published_books, n = 5, with_ties = FALSE)
```







