---
title: "R Notebook"
output: html_notebook
---



#### MVP Task 1

```{r}
library(tidyverse)
library(janitor)
library(e1071)
library(infer)

ames <- read_csv("data/ames.csv") %>% 
  clean_names() %>% 
  glimpse()
```

#### Task 2

```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(aes(y = ..density..), col = "white", fill = "steel blue", alpha = 0.7) + 
  stat_function(
    fun = dnorm, 
    args = list(
      mean = mean(ames$lot_area), 
      sd = sd(ames$lot_area)
    ), 
    col = "red"
  )
```

The distribution looks very skewed. Let's investigate. 

```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_boxplot()
```
```{r}
ames %>% 
  summarise(skewness = skewness(lot_area, type = 1))
```
The box plot shows the number of lot_area values in the 4th quartile of the distribution which are large in value. This is also reflected in the skewness value for the data which is very highly right skewed. 

#### Task 3.

Bootstrap sample of the full population of `ames` data, 5000 times.

```{r}
bootstrap_ames <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

bootstrap_ames
```



Histogram of the bootstrap data set:

```{r}
bootstrap_ames %>% 
  ggplot() +
  aes(x = stat) + 
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7)
```

#### Task 4. 

95% confidence interval for mean(lot_area)
 
```{r}
bs_ames_ci_95 <- bootstrap_ames %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

bs_ames_ci_95
```

#### Task 5.

A 99% CI is broader than the 95% CI.

```{r}
bs_ames_ci_99 <- bootstrap_ames %>% 
  get_confidence_interval(level = 0.99, type = "percentile")

bs_ames_ci_99
```

#### Task 6.

Mean lot area from bootstrap sample.

```{r}
mean_bs_lot_area <- bootstrap_ames %>% 
  summarise(mean = mean(stat))

mean_bs_lot_area
```
Mean lot area from the original ames data set.

```{r}
mean_lot_area <- ames %>% 
  summarise(mean = mean(lot_area))

mean_lot_area
```









