---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(modelr)

library(tidyverse)
titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```


## Question 1.1

```{r}
titanic_clean <- titanic_set %>% 
  drop_na() %>% 
  mutate(sex = as.factor(sex), 
         survived = factor(survived, levels = c(0, 1), labels = c("No", "Yes")), 
         class = factor(pclass, levels = c(1, 2, 3), labels = c("Upper", "Middle", "Lower")), 
         embarked = as.factor(embarked),
         age_status = as.factor(if_else(age <= 16, "child", "adult"))
         ) %>%
  select(sex, survived, class, embarked, age_status, sib_sp, parch)
```


## Question 1.2

```{r}
titanic_clean %>% 
  ggplot() +
  aes(x = class, fill = survived) +
  geom_bar(position = "dodge") + 
  labs(
    title = "Passengers by class and survival status"
  )
```

```{r}
titanic_clean %>% 
  ggplot() +
  aes(x = embarked, fill = survived) +
  geom_bar(position = "dodge") + 
  labs(
    title = "Passengers by embarcation point and survival status"
  )
```


```{r}
titanic_clean %>% 
  ggplot() +
  aes(x = sex, fill = survived) +
  geom_bar(position = "dodge") + 
  labs(
    title = "Passengers by gender and survival status"
  )
```

Many more of the wealthier survived. 

More of those joining in Cherbourg and Southampton survived

More women survived than men. 

Variables of use: class, sex, embarcation point


## Question 3

```{r}
# number of rows in dataset
n_data <- nrow(titanic_set)

# create a test sample index
test_index <- sample(1:n_data, size = n_data * 0.2)

# create test set
titanic_test  <- slice(titanic_clean, test_index)

# create training set
titanic_train <- slice(titanic_clean, -test_index)
```

```{r}
# survival split in test set
titanic_test %>% 
  tabyl(survived)
```

```{r}
# survival split in train set
titanic_train %>% 
  tabyl(survived)
```

The split is fairly even!


## Question 4


```{r}
titanic_fit <- rpart(
  formula = survived ~ .,
  data = titanic_train, 
  method = "class"
)

rpart.plot(titanic_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4)
```

## Question 5

The top node indicates that the most likely outcome for passengers is `survival`. Yeah!
Almost 68% of passengers survived from this data sample. 

Moving down to the second level (below the root node):
If you were female, you were likely to survive (94.29%). Females accounted for 48% of the total passengers. 
If you were male, you only had a 43.42% chance of survival despite making up 52% of the total passengers. 

Moving down again:
If you were male and embarked in Queenstown or Southampton, you had a 60% chance of surviving (1 - 40%) and made up 34% of total passengers. 
If you were male and embarked in Cherbourg, your chances of survival were 50/50 and you accounted for 18% of passengers.

For those males from Cherbourg with family, you were more likely to survive (62.5%) than if you had no family with you (44.44%). 


# Question 6

```{r}
# add predictions
titantic_test_pred <- titanic_test %>% 
  add_predictions(titanic_fit, type = "class")

titantic_test_pred
```

```{r}
library(yardstick)

conf_mat <- titantic_test_pred %>% 
  conf_mat(truth = survived, estimate = pred)

conf_mat
```

The model correctly predicted that 11 passengers died and that 17 passengers survived. 

The model incorrectly predicted that 7 passengers died when in fact they survived (false negative). 

The model incorrectly predicted that 2 passengers survived when in fact they died (false positive). 


```{r}
accuracy <- titantic_test_pred %>% 
  accuracy(truth = survived, estimate = pred)

accuracy
```
The model is 76% accurate.











