---
title: "week_10_day_01 homework"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
```

```{r}
# load in the data
project <- read_csv("data/project_management.csv")
```

```{r}
# plot project data
project %>% 
  ggplot() +
  aes(x = estimated_length, y = actual_length) +
  geom_point()
```

```{r}
# calculate correlation coefficient
project %>% 
  summarise(cor(estimated_length, actual_length))
```
There is a very strong positive correlation between the two variables. 


```{r}
# simple linear regression of project variables
project_model <- lm(formula = actual_length ~ estimated_length, data = project)

project_model
```
The regression coefficient is 1.223, meaning that for every 1 unit increase in estimated length, the actual length increases by 1.223


```{r}
# correlation coefficients
glance_output <- clean_names(glance(project_model))

glance_output
```
A r^2 value of 0.6474 means that 64.74% of the variation in `actual length` can be explained by variation in `estimated length`.


```{r}
library(ggfortify)
```

```{r}
# diagnostic plots
autoplot(project_model)
```

The p_value is less than a 0.01 significance level so would imply that the coefficient of determination (i.e. r^2) is significantly different from zero. 

This conclusion is subject to interpretation of the above diagnostic plots.

- Residuals vs Fitted: 
This plot tests the independence of the residuals. 
Ideally the residuals would be randomly scattered around zero. This is the case in the above plot.

- Normal Q-Q:
This plot tests the normality of the residuals.
All points should lie close to the line. In the above plot they do, indicating the residuals to be normally distributed. 

- Scale-Location:
This plot tests the constancy of variation of the residuals. 
Ideally the residuals should all lie in a band of ficed width above the x-axis.
The blue line in the graph above stays close to a constant value so is fine. 

**Conclusion:**
Given that the diagnostic plots indicate that the residuals are:
- independent of each other
- normally distributed, and 
- are fairly constant in variation across the range of data
we can conclude that the low p-value indicates that the relationship between estimated length and actual length is statistically significant.  



