---
title: "Week_05_day_02 homework"
author: "Malcolm Speight"
date: "2022-12-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


<br/> 

##### **Question 1**

```{r, include = FALSE}
library(tidyverse)
library(CodeClanData)
```


##### **Question 2**

The distribution shows that the most common backpack weight is 10 with only 1 backpack weighing more than 30. 

```{r}
ggplot(backpack) +
  aes(x = backpack_weight) +
  geom_bar() +
  scale_x_continuous(breaks = 1:40) +
  scale_y_continuous(breaks = 1:15) +
  labs(
    x = "\nweight", 
    y = "count", 
    title = "Backpack weights\n"
  )
```

<br/> 

##### **Question 3**

The most common backpack weight for women is 10 whilst for men it is 11. 

The heaviest backpack carried by a woman is 25 whilst one man carried a backpack of weight 35. 

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, fill = sex) +
  geom_bar(position = "dodge") +
  scale_x_continuous(breaks = 1:40) +
  scale_y_continuous(breaks = 1:15) +
  labs(
    x = "\nweight", 
    y = "count", 
    title = "Backpack Weights by Gender\n", 
    fill = ""
  )
```


<br/> 

##### **Question 4**

There doesn't seem to be much of a relationship between body weight and backpack weight.

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, y = body_weight) +
  geom_point() +
  labs(
    x = "\nbackpack weight", 
    y = "body weight\n", 
    title = "Backpack Weights by Body Weight\n"
    )
```

<br/> 

##### **Question 5**

It would appear that women are carrying heavier backpacks in relation to their body weight than men are. 

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, y = body_weight, colour = sex) +
  geom_point() +
  labs(
    x = "\nbackpack weight", 
    y = "body weight\n", 
    title = "Backpack Weights by Body Weight and Gender\n"
    )
```


<br/> 

##### **Question 6**

About a third of students report back pain. 

```{r}
backpack %>% 
  mutate(back_problems_adj = ifelse(back_problems, TRUE, FALSE)) %>% 
  mutate(back_problems_adj = factor(back_problems_adj, levels = c("TRUE", "FALSE"))) %>% 
  ggplot() +
  aes(x = back_problems_adj) +
  geom_bar() +
  
  
  labs(
    x = "\nback pain", 
    y = "count\n",
    title = "Back problems in students\n"
  )
```

<br/> 


##### **Question 7**

About 20% of male students report back pain. That ratio is closer to 50% for female students. 

```{r}
backpack %>% 
  mutate(back_problems_adj = ifelse(back_problems, TRUE, FALSE)) %>% 
  mutate(back_problems_adj = factor(back_problems_adj, levels = c("TRUE", "FALSE"))) %>% 
  ggplot() +
  aes(x = back_problems_adj, fill = sex) +
  geom_bar(position = "dodge") +
  scale_y_continuous(breaks = c(10, 20, 30, 40)) +
  labs(
    x = "\nback pain", 
    y = "count\n",
    title = "Back problems in students\n", 
    fill = ""
  )
```


<br/> 

##### **Question 8**

Those students suffering back pain (both male and female) tend to have heavier backpacks compared to 
students not suffering back pain. 

```{r}
backpack %>% 
  mutate(back_problems_adj = ifelse(back_problems, TRUE, FALSE)) %>% 
  mutate(back_problems_adj = factor(back_problems_adj, levels = c("TRUE", "FALSE"))) %>% 
  ggplot() +
  aes(x = body_weight, y = backpack_weight, colour = sex) +
  geom_point() +
  facet_grid(back_problems_adj ~ sex) +
  labs(
    x = "\nbody weight",
    y = "backpack weight\n", 
    title = "Back Problems by Backpack Weight and Gender\n"
    )
```

