---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(mosaicData)
library(janitor)
library(GGally)
```

### Task 1

```{r}
housing_prices <- read_csv("data/housing_prices.csv") %>% 
  clean_names()

summary(housing_prices)
```
Looking at the Kaggle website, the `median_income` variable is measured in tens-of-thousands of US Dollars, whilst the `median_house_value` is measured in US Dollars. 

Convert `median_income` to US Dollars. 

```{r}
housing_prices_scaled <- housing_prices %>% 
  mutate(median_income = median_income * 10000)
```

Variable `total_bedrooms` has 200 missing values. Replace with mean of `total_bedrooms`.

```{r}
housing_prices_scaled <- housing_prices_scaled %>% 
  mutate(total_bedrooms = coalesce(total_bedrooms, mean(total_bedrooms, na.rm = TRUE)))
```

Check for aliases. 

```{r}
alias(lm(total_rooms ~., data = housing_prices_scaled))
```
No aliases found. 


### Task 2

```{r}
ggpairs(housing_prices_scaled)
```

Correlation between `total_rooms` and `total_bedrooms` is 0.931. 


### Task 3

```{r}
# remove total_bedrooms from dataset
housing_prices_scaled <- housing_prices_scaled %>% 
  select(-total_bedrooms)
```


### Task 4

```{r}
ggpairs(housing_prices_scaled)
```

Looks like there are 3 key variables for `median_house_value`: `latitude`, `total_rooms` and `median_income`

Let's plot each of these against `median_house_value`.

```{r}
housing_prices_scaled %>% 
  ggplot() + 
  aes(x = latitude, y = median_house_value) + 
  geom_point() +
  scale_y_continuous(labels = scales::comma)
```

```{r}
housing_prices_scaled %>% 
  ggplot() + 
  aes(x = total_rooms, y = median_house_value) + 
  geom_point() + 
  scale_y_continuous(labels = scales::comma)
```
`total_rooms` is the number of rooms in a block. Divide this by number of households to get average rooms per household:

```{r}
housing_prices_scaled <- housing_prices_scaled %>% 
  mutate(avg_rooms_per_house = total_rooms / households)
```

```{r}
housing_prices_scaled %>% 
  ggplot() + 
  aes(x = avg_rooms_per_house, y = median_house_value) + 
  geom_point() +
  scale_y_continuous(labels = scales::comma)
```
Plenty of outliers in the `avg_rooms_per_house` variable



```{r}
housing_prices_scaled %>% 
  ggplot() + 
  aes(x = median_income, y = median_house_value) + 
  geom_point() +
  scale_y_continuous(labels = scales::comma)
```


### Task 5

```{r}
housing_prices_scaled %>% 
  distinct(ocean_proximity)
```
There are 5 unique values/levels within `ocean_proximity`, so that means you require n - 1 dummy variables, which is 4. 


### Task 6

```{r}
library(ggfortify)

house_model <- lm(median_house_value ~ median_income, data = housing_prices_scaled)

autoplot(house_model) # diagnostics
```
residuals vs fitted: looks ok with residuals randomly scattered around zero
normal q-q: looks ok-ish but outliers at the low and high end distort the normal distribution
scale-location: definite funnelling effect

```{r}
summary(house_model)
```
**Conclusion**
house_model produces a r^2 value of 0.4129 with a statistically significant p-value. 


### Task 7

Add `total_rooms` to the model. 

```{r}
housing_prices_scaled %>% 
  summarise(cor = cor(median_house_value, total_rooms))
```
A very weak positive correlation between the two variables. Great!

Nevertheless, add it to the model:

```{r}
house_model_2 <- lm(median_house_value ~ median_income + total_rooms,
                    data = housing_prices_scaled)
```

```{r}
# diagnostics
autoplot(house_model_2)
```
 Very similar to the first model. 

residuals vs fitted: looks ok with residuals randomly scattered around zero
normal q-q: looks ok-ish but outliers at the low and high end distort the normal distribution
scale-location: definite funnelling effect

```{r}
summary(house_model_2)
```
**Conclusion:**
house_model_2 does not change let alone improve on the r^2 value of 0.4129 from the first house_model. 
Indeed, the p-value for `total_rooms` indicates that as a predictor it is not statistically significant. 













