---
title: "day_02 homework"
output: html_notebook
---

# Question 1

```{r}
CodeClanData::drinks_content
```
# dimensions, rows then cols
```{r}
dim(drinks_content)
```

# variable names
```{r}
names(drinks_content)
```

# variable type
```{r}
glimpse(drinks_content)
```

# Question 2

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  filter(!is.na(Calories))
```

# Question 3
# assuming a carry over from Q2

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  filter(!is.na(Calories)) %>% 
  mutate(calorie_diff = Calories - 135)
```

# Question 4

```{r}
drinks_content %>% 
  group_by(Beverage_category) %>% 
  summarise(Mean_calories = mean(Calories, na.rm = TRUE)) %>% 
  arrange(desc(Mean_calories))
```

# Most calories: Smoothies, Frappuccino Blended Coffee, Signature Espresso Drinks

# Least calories: Classic Espresso Drinks, Shaken Iced Beverages, Coffee


# Question 5

```{r}
drinks_content %>%
  group_by(Beverage_prep) %>% 
  summarise(Avg_calories = mean(Calories, na.rm = TRUE)) %>% 
  arrange(desc(Avg_calories))
```

# Question 6
# assume question refers to average Beverage_prep calories

# calculate average calories in all drinks
```{r}
average_cal_all_drinks <- drinks_content %>%
  summarise(avg_cal = mean(Calories, na.rm = TRUE)) %>% 
  pull()
```

# print average all-drinks cal value
```{r}
average_cal_all_drinks
```

# Beverage prep types with greater than the average all-drinks cal value 
```{r}
drinks_content %>% 
  group_by(Beverage_prep) %>% 
  summarise(Avg_calories = mean(Calories, na.rm = TRUE)) %>% 
  filter(Avg_calories > average_cal_all_drinks) %>% 
  arrange(desc(Avg_calories))
```

# Question 7

# Best coffee drink for those watching the calories

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  filter(Beverage_category == "Coffee") %>% 
  arrange(Calories) %>% 
  head(1)
```
